A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium

### 2020-10-17 simplified_sdc_intro/paper.tex
some updates
more updates on SDC method

@@

@@

We use the following conditions to determine if we are in NSE:
\begin{align}
\rho &> \rho_\mathrm{nse} \\
T &> T_\mathrm{nse} \\
X(\isotm{C}{12}) &< A_\mathrm{nse} \\
X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
\end{align}
where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.

@@

@@

To test the coupling of the NSE table to the hydrodynamics, we run a
similar problem as above, except with the thermodynamic conditions
appropriate for the matter to be in NSE.

>>>>

\newcommand{\rhonse}{{\rho_\mathrm{nse}}}
\newcommand{\tnse}{{T_\mathrm{nse}}}
\newcommand{\Anse}{{A_\mathrm{nse}}}
\newcommand{\Bnse}{{B_\mathrm{nse}}}

@@

\label{sec:initialization}

@@

We treat a zone as being in NSE if the density and temperature exceed
some threshold and the composition is mainly $\alpha$ and Fe-group
nuclei, where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52},
\isot{Fe}{54}, and \isot{Ni}{56}.  The full condition is:
\begin{align}
\rho &> \rhonse \\
T &> \tnse \\
X(\isotm{C}{12}) &< \Anse \\
X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse
\end{align}
Typical values are $\rhonse = 3\times 10^8~\gcc$, $T_\mathrm{nse} =
3\times 10^9~\mathrm{K}$, and $\Anse = 0.12$ and $\Bnse=
0.88$---these are the values used in \citet{ma:2013}.

@@

With SDC evolution, when we are in NSE, we need to do the advective
and reactive updates together.  With the tabulated NSE reaction
values, we get instantaneous values of the reaction sources, but the only
time derivative we have is for $Y_e$.  We will
use a basic predictor-corrector update:
\begin{equation}
\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \Rb (\Uc^n)
\end{equation}
where $\Rb(\Uc^n)$ is the reaction source vector evaluated by calling the NSE
table with the time-level $n$ state.
In particular, this will also operate on the auxiliary composition variables,
including $Y_e$:
\begin{equation}
(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Adv{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
\end{equation}
From this, we get $\rho^\star$, $(Y_e)^\star$, $(X_k)^\star$, and
$e^\star$.  In order call the get the temperature, we need to call the
EOS which requires $\bar{A}^\star$.  Since we cannot use the NSE table
without the temperature, we need to

@@

To test the coupling of the NSE table to the hydrodynamics, we run a
similar problem as above, except with the thermodynamic conditions
appropriate for the matter to be in NSE.  The initial conditions are:
\begin{align}
\rho &= \rho_0 \\
T &= T_0 \left [ 1 + (\delta T) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] \\
Y_e &= (Y_e)_0 \left [ 1 + (\delta Y_e) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] 
\end{align}
where $r$ is the distance from the center of the domain and the
remaining parameters are listed in Table~\ref{table:nse}.  This has
$Y_e$ varying between $[0.47, 0.5]$ initially\MarginPar{check}.  The mass fractions
and remaining composition variables are then initialized from the NSE table, as described in section~\ref{sec:initialization}.

The domain has a size $[0, L]^2$, and the timestep is fixed as:
\begin{equation}
\Delta t = 5\times 10^{-4} \left ( \frac{128}{N} \right )~\mathrm{s}
\end{equation}
where $N$ is the number of zones in each direction.

\begin{deluxetable}{lcc}
\tablecaption{\label{table:nse} NSE convergence test problem.}
\tablehead{\colhead{parameter} & \colhead{value}}
\startdata
$\rho_0$ & $5\times 10^8~\gcc$ \\
$T_0$    & $4\times 10^9~\mathrm{K}$ \\
$(\delta T)$ & $0.2$ \\
$(Y_e)_0$   & $0.5$ \\
$(\delta Y_e)$ & $-0.05$ \\
$\lambda$   & $2\times 10^7~\mathrm{cm}$ \\
$L$         & $10^8~\mathrm{cm}$ \\
\enddata
\end{deluxetable}
